---
title: "Replication Analysis: AI Access, Resilience, Motivation, and Academic Outcomes"
author: "Replication Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 6, fig.height = 4)
set.seed(1234)
```

# Introduction

This report replicates the empirical analysis examining whether AI access
(*Access_Plus*) predicts psychological resilience (*Resilience*), whether
motivation (*Motivation*) mediates this relationship, and whether institutional
barriers (*Bank_Barrier*) and academic performance (*GPA*) relate to these
outcomes.

# Method

## Participants and Measures

Data were drawn from `AI_Resilience_Empirical_Dataset.xlsx` (*N* = 60). All
scales were rated on 1-5 Likert-type scales; GPA ranged from 80 to 99.

## Analytic Strategy

Analyses included descriptive statistics, Pearson correlations, OLS regression,
assumption checks (Shapiro-Wilk, NCV test, VIF), a one-way ANOVA by status
group, and a bootstrap mediation analysis (1,000 resamples; Tingley et al., 2014).

# Results

```{r load}
library(readxl); library(tidyverse); library(psych); library(car)
library(mediation); library(ggplot2); library(gridExtra)
dat <- read_excel("AI_Resilience_Empirical_Dataset.xlsx") %>%
  mutate(across(where(is.numeric), as.numeric))
```

## Descriptive Statistics

```{r descriptives}
describe(dat[, c("Access_Plus","Bank_Barrier","Resilience","Motivation","GPA")])
```

Descriptive statistics indicated moderate levels of AI access
(*M* = `r round(mean(dat$Access_Plus), 2)`, *SD* = `r round(sd(dat$Access_Plus), 2)`)
and resilience (*M* = `r round(mean(dat$Resilience), 2)`, *SD* = `r round(sd(dat$Resilience), 2)`).

## Correlations

```{r correlations}
round(cor(dat %>% select(-ID, -Status), use = "pairwise"), 3)
```

## Regression Analyses

```{r regressions}
m1 <- lm(Resilience ~ Access_Plus, data = dat)
m2 <- lm(Motivation  ~ Access_Plus, data = dat)
m3 <- lm(Resilience ~ Access_Plus + Motivation + Bank_Barrier, data = dat)
summary(m1); summary(m2); summary(m3)
car::vif(m3)
```

The full model tested whether AI access and motivation predicted resilience
above institutional barriers. Unstandardized coefficients (*b*), standard
errors, *t* statistics, and *p* values are reported in APA style from the
output above.

## Assumption Checks

```{r assumptions}
shapiro.test(residuals(m3))
car::ncvTest(m3)
```

## ANOVA by Status Group

```{r anova}
dat$Status_f <- factor(dat$Status, levels = c(1, 2),
                       labels = c("Group1", "Group2"))
fit_aov <- aov(Resilience ~ Status_f, data = dat)
summary(fit_aov)
s <- summary(fit_aov)[[1]]
eta2 <- s[1, "Sum Sq"] / (s[1, "Sum Sq"] + s[2, "Sum Sq"])
cat("eta-squared =", round(eta2, 3), "\n")
```

## Mediation Analysis

```{r mediation, cache=TRUE}
set.seed(1234)
med_fit <- mediate(m2, m3, treat = "Access_Plus", mediator = "Motivation",
                   boot = TRUE, sims = 1000)
summary(med_fit)
plot(med_fit)
```

The indirect effect of AI access on resilience through motivation was tested
with 1,000 bootstrap resamples. Report the ACME, average direct effect (ADE),
total effect, and proportion mediated from the output above; 95% confidence
intervals excluding zero indicate significant mediation.

## Figures

```{r figures}
p1 <- ggplot(dat, aes(Access_Plus, Resilience)) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  geom_smooth(method = "lm") +
  labs(x = "AI Access", y = "Resilience") + theme_minimal()
p2 <- ggplot(dat, aes(Motivation, Resilience, color = Status_f)) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Motivation", y = "Resilience", color = "Status") + theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```

# Discussion

Results indicate that AI access was positively associated with resilience,
with motivation examined as a potential mediator and institutional barriers as
a covariate. Given the modest sample (*N* = 60) and single-item measures,
findings should be interpreted cautiously and replicated with larger,
multi-item samples.

# References

- Tingley, D., Yamamoto, T., Hirose, K., Keele, L., & Imai, K. (2014).
  *mediation*: R package for causal mediation analysis. *Journal of
  Statistical Software, 59*(5).
- Imai, K., Keele, L., & Tingley, D. (2010). A general approach to causal
  mediation analysis. *Psychological Methods, 15*(4), 309-334.
- R Core Team. (2024). *R: A language and environment for statistical
  computing*.
